Tight Bounds for Online Balanced Graph Partitioning

### 2020-05-08 paper/disc20clustering.tex
Wip
intro for rematching

 \newcommand{\OPT}{\textsf{OPT}\xspace}
  \newcommand{\OPTM}{\mathit{OPT}}
 \newcommand{\ALG}{\textsf{ALG}\xspace}
 \newcommand{\PPL}{\textsf{PPL}\xspace}
 \newcommand{\OBRP}{BRP}
 \newcommand{\PPOBRP}{PP-BRP}
 \newcommand{\dist}{\textsf{dist}}
 \newcommand{\TAlg}{{\ensuremath{\textsf{ALG}_{3}}}\xspace}
  \newcommand{\RM}{\textsf{RM}\xspace} % rematching alg

@@

\subsection{Improved Algorithm for Online Rematching}
\label{sec:k2}

\maciek{Intro: compare to previous 7-competitive algorithm. Note strict competitive ratio vs large additive term. Also, emphasize that in this section the algorithm is different than DET2}

@@

\textbf{The charging scheme.}
Consider two  pairs that share a same node, 
i.e.~\emph{intersecting pairs},
and the set of requests that causes the collocation of both pairs,
at some  times during  $[1,m]$.
%i.e.~the entire duration of the sequence.
Observe that \OPT must pay a non-zero cost
for these requests over the entire $\sigma$,
since it cannot have both pairs initially collocated.
%More precisely,
%OPT  pays a total cost at least
%	$\min{ \{ 2\lambda, \lambda+\alpha, 2\alpha \} }$ for, respectively,
%	serving requests to both pairs remotely,
%	serving requests to one pair remotely and collocating the other,
%	or collocating both pairs. 
%
%for serving (some of) these requests remotely or (and) collocating (one of) the pairs,
%regardless of its choices.
%Then the cost per pair, in particular the cost w.r.t.~$\{u,v\}$, is
%$\mathit{OPT} (\sigma_{\{u,v\}}) \geq c_{R_t} / 2$.
%
However,
we can charge the whole cost to \OPT only the first time \RM collocates a pair,
and not at any consequent time when \RM collocates it a second time.
Otherwise,
 \OPT is possibly charged for a same cost repeatedly.
For this reason,
we charge \OPT a cost inflicted by a pair,
if and only if  \OPT incurs the cost some time since the last time \RM separated the pair.
%%

\begin{proof}[Proof of Theorem \ref{thm:k=2}]
	Assume on request $\sigma_{t}=\set{u,v}$, 
	i.e., at (time) $t$,
	Alg collocates the pair $\set{u,v}$.
	This means,
	the value of $C_{\set{u,v}}$ at $t$,
	denoted by $C^t_{\set{u,v}}$, 
	reaches $\lambda$ immediately before \RM resets the counter.
	I.e.,
	$ C^t_{\set{u,v}} = \lambda$.
	%	In case the first collocation of $\set{u,v}$ occurs at $t$,
	%	if the pair is collocated in the initial configuration then
	%	we set $t''=0$,
	%	else,
	%	we set $ t'' = -\infty, t'=0$.
	We denote the set of all requests to a pair $\set{x,y}$ that arrive
	during $[t_1,t_2]$ by $\sigma_{\set{x,y}}[t_1,t_2]$,
	and its overall cost to \OPT by
	$\mathit{OPT} (\sigma_{\{u,v\}}[t_1,t_2])$.
	We may use $\sigma_{\set{x,y}}$ whenever
	 the interval $[t_1,t_2]$ is clear from the context.
	%
	%Else, this is not the first time and
	%Alg collocates $\{u,v\}$ also some time prior to $t$.

@@

	First,
	we bound  costs incurred by \RM for requests that
	lead to the collocation of $\{u,v\}$ at time $t \in T$, where
	$T := \{ i \in [1,m] ~\vert~ \exists \{x,y\}: C^{i}_{\{x,y\}} = \lambda \}$
	is the set of times when \RM performs a collocation.
	By definitions of $t$ and $t'$,
	the overall cost of requests in $\sigma_{\set{u,v}}$ incurred by \RM,
	i.e., the total cost of remote serving
	and the moving cost, is
	$\lambda + 2\alpha$.	
	Next,
	we bound costs incurred by \RM
	for requests that do not lead to collocations until the  end of the sequence at $t=m$.
	Assume $\{u,v\}$ is not collocated at $t=m$
	and $0 < C^{m}_{ \{u,v\} } < \lambda $,
	which means \RM pays $C^{m}_{ \{u,v\} }$
	for  requests in $\sigma_{\set{u,v}}(t',m]$.
    Then the overall cost to \RM is
$	\mathit{RM} (\sigma)
=
\sum_{ t \in T}(\lambda + 2\alpha) +
\sum_{\{u,v\}} C^{m}_{\{u,v\}}	
$.
%	\begin{align*}	%\label{eq:costAlg}
%	\mathit{RM} (\sigma)
%	=
%	\sum_{ t \in T}(\lambda + 2\alpha) +
%	\sum_{\{u,v\}} C^{m}_{\{u,v\}}	
%	\leq
%	\sum_{ t \in T}(\lambda + 2\alpha) +
%	\sum_{\{u,v\}} C^{m}_{\{u,v\}}	
%	=
%	\sum_{ t \in T} 3\alpha +
%	\sum_{\{u,v\}} C^{m}_{\{u,v\}}
%	.
%	\end{align*}

	Next,
	we bound  costs incurred by \OPT for requests that trigger collocation of $\{u,v\}$ at $t \in T$.
	If $t$ is the first time that \RM collocates $\{u,v\}$ then  \OPT pays
	$\lambda$ for serving requests in $\sigma_{\set{u,v}}[0,t]$ (remotely),
	or $\alpha$ for collocating the pair and
	serving (some of) the requests with  cost zero.
	Therefore in this case,
	$\OPTM (\sigma_{\{u,v\}}(0,t]) \geq  \min{\{ \lambda,\alpha \}}$.
	%
	Otherwise,
	it is not the first collocation and
	consider times $t'$ and $t''$ as define previously,
	 and let 
	$R_t := \sigma_{\{u,w\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t] $.
	We define $R_{t'}$ for the collocation at $t'$  analogously.
	Then,
	$\OPTM(R_t) = \mathit{OPT} (\sigma_{\{u,w\}}) 
	+ \OPTM(\sigma_{\{u,v\}}) $.
	Note that \OPT cannot have both pairs collocated at the same time.
	However, it can have one of the pairs already collocated prior to its respective interval.
	Let us assume  \OPT has one pair,
	e.g.~$\{u,v\}$ collocated already prior its respective interval, i.e.~$(t',t]$,
	 and keeps it so during the interval.
	 Then it pays zero while serving $\sigma_{\{u,v\}}$.
	Hence,
	it must pay $\alpha$ for collocating the other pair, in this case $\{u,w\}$,
	or (resp., and) it pays (resp., up to) $\lambda$ for serving (resp., some of) requests in $\sigma_{\{u,w\}}$. 
	Therefore in any case,
	$\OPTM(R_t) \geq  \min{\{ \lambda,\alpha \}}$.
	\begin{comment}
	We distinguish three (non-exhaustive) cases for the cost to \OPT w.r.t.~$R_t$.
	Latter,
	 we argue that \OPT cannot save cost by deviating from these cases.
	\begin{enumerate}[label=\roman*.]
		\item
		\OPT collocates $\{u,w\}$ at  $t''$, $\{u,v\}$  at $t'$,
		and keeps them so during $(t'',t']$ and $(t',t]$, respectively.
		Therefore it serves all requests in $R_t$ with no additional cost and
		let $\OPTM_1 (R_t) := 2\alpha$.
		\item
		\OPT has one of the pairs  already collocated and the other pair separated,
			  prior to and during their respective intervals.
		Hence,
		it pays only $\lambda$ for serving all requests in $R_t$
		and let $\OPTM_2 (R_t) := \lambda$.
		\item
		\OPT has one of the pairs already collocated prior to its interval,
		and it collocates the other pair during the respective interval.
		Then it pays 0 for one pair and $\alpha$ for the other and 
		let $\OPTM_3 (R_t) := \alpha$.
	\end{enumerate}
%	We consider cases (ii) and (iii) also for the pair $\{u,w\}$, symmetrically.
	If \OPT behaves in one of these cases,
	then
	$\mathit{OPT} (R_t) \geq
		\min{\{ \OPTM_1(R_t), \OPTM_2(R_t), \OPTM_3(R_t) \}}  =
		\min{ \{ \lambda, \alpha \}}$.
	Otherwise,
	\OPT deviates from these cases,
	e.g.~by collocating a pair after paying for some of requests to this pair.
	However,
	no deviation can decrease the cost to \OPT. \mahmoud{prove this!}
	Therefore in any case,
	$\mathit{OPT} (R_t) \geq  \min{ \{ \lambda, \alpha \}}$.
	\end{comment}

>>>>

 \newcommand{\OPT}{\textsf{OPT}\xspace}
  \newcommand{\OPTM}{\mathit{OPT}}
 \newcommand{\ALG}{\textsf{ALG}\xspace}
 \newcommand{\PPL}{\textsf{PPL}\xspace}
 \newcommand{\OBRP}{BRP\xspace}
 \newcommand{\PPOBRP}{PP-BRP}
 \newcommand{\dist}{\textsf{dist}}
 \newcommand{\TAlg}{{\ensuremath{\textsf{ALG}_{3}}}\xspace}
  \newcommand{\RM}{\textsf{RM}\xspace} % rematching alg

@@

\subsection{Improved algorithm for Online Rematching} \label{sec:k2}
In this section,
we present \RM,
 an algorithm for the \OBRP restricted to clusters of capacity 2,
i.e.~$k=2$.
We interpret a pair of nodes collocated in one cluster as a matched pair.
Hence,
the problem is an online variant of the classic matching problem where
a matched pair can separate from each other in order to ``rematch'' with two other nodes.
This is known as the  \emph{Online Rematching} 
problem and a (non-strict) 7-competitive algorithm is already given by \cite{repartition-disc},
in which the ratio comes with an additive factor in $O(\alpha\ell^2)$.
We do not only improve upon their competitive ratio,
but also show that our ratio holds \emph{strictly}
(i.e., with no additive factor).
In addition,
our algorithm is similar
and slightly simpler than the one in \cite{repartition-disc}, 
while our analysis is significantly more simple and concise,
thanks to the charging scheme we devise here.

@@

\textbf{The charging scheme.}
Consider two  pairs that share a same node, 
i.e.~\emph{intersecting pairs},
and the set of requests that causes the collocation of both pairs,
at some  times during  $[1,m]$.
%i.e.~the entire duration of the sequence.
Observe that \OPT must pay a non-zero cost
for these requests over the entire $\sigma$,
since it cannot have both pairs initially collocated.
However,
we can charge the whole cost to \OPT only the first time \RM collocates a pair,
and not at any consequent time when \RM collocates it a second time.
Otherwise,
 \OPT is possibly charged for a same cost repeatedly.
For this reason,
we charge \OPT a cost inflicted by a pair,
if and only if  \OPT incurs the cost some time since the last time \RM separated the pair.
%%

\begin{proof}[Proof of Theorem \ref{thm:k=2}]
	Assume on request $\sigma_{t}=\set{u,v}$, 
	i.e., at (time) $t$,
	Alg collocates the pair $\set{u,v}$.
	This means,
	the value of $C_{\set{u,v}}$ at $t$,
	denoted by $C^t_{\set{u,v}}$, 
	reaches $\lambda$ immediately before \RM resets the counter.
	I.e.,
	$ C^t_{\set{u,v}} = \lambda$.
	We denote the set of all requests to a pair $\set{x,y}$ that arrive
	during $[t_1,t_2]$ by $\sigma_{\set{x,y}}[t_1,t_2]$,
	and its overall cost to \OPT by
	$\mathit{OPT} (\sigma_{\{u,v\}}[t_1,t_2])$.
	We may use $\sigma_{\set{x,y}}$ whenever
	 the interval $[t_1,t_2]$ is clear from the context.
	%
	%Else, this is not the first time and
	%Alg collocates $\{u,v\}$ also some time prior to $t$.

@@

	First,
	we bound  costs incurred by \RM for requests that
	lead to the collocation of $\{u,v\}$ at time $t \in T$, where
	$T := \{ i \in [1,m] ~\vert~ \exists \{x,y\}: C^{i}_{\{x,y\}} = \lambda \}$
	is the set of times when \RM performs a collocation.
	By definitions of $t$ and $t'$,
	the overall cost of requests in $\sigma_{\set{u,v}}$ incurred by \RM,
	i.e., the total cost of remote serving
	and the moving cost, is
	$\lambda + 2\alpha$.	
	Next,
	we bound costs incurred by \RM
	for requests that do not lead to collocations until the  end of the sequence at $t=m$.
	Assume $\{u,v\}$ is not collocated at $t=m$
	and $0 < C^{m}_{ \{u,v\} } < \lambda $,
	which means \RM pays $C^{m}_{ \{u,v\} }$
	for  requests in $\sigma_{\set{u,v}}(t',m]$.
    Then the overall cost to \RM is
$	\mathit{RM} (\sigma)
=
\sum_{ t \in T}(\lambda + 2\alpha) +
\sum_{\{u,v\}} C^{m}_{\{u,v\}}	
$.

	Next,
	we bound  costs incurred by \OPT for requests that trigger collocation of $\{u,v\}$ at $t \in T$.
	If $t$ is the first time that \RM collocates $\{u,v\}$ then  \OPT pays
	$\lambda$ for serving requests in $\sigma_{\set{u,v}}[0,t]$ (remotely),
	or $\alpha$ for collocating the pair and
	serving (some of) the requests with  cost zero.
	Therefore in this case,
	$\OPTM (\sigma_{\{u,v\}}(0,t]) \geq  \min{\{ \lambda,\alpha \}}$.
	%
	Otherwise,
	it is not the first collocation and
	consider times $t'$ and $t''$ as define previously,
	 and let 
	$R_t := \sigma_{\{u,w\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t] $.
	We define $R_{t'}$ for the collocation at $t'$  analogously.
	Then,
	$\OPTM(R_t) = \mathit{OPT} (\sigma_{\{u,w\}}) 
	+ \OPTM(\sigma_{\{u,v\}}) $.
	Note that \OPT cannot have both pairs collocated at the same time.
	However, it can have one of the pairs already collocated prior to its respective interval.
	Let us assume  \OPT has one pair,
	e.g.~$\{u,v\}$ collocated already prior its respective interval, i.e.~$(t',t]$,
	 and keeps it so during the interval.
	 Then it pays zero while serving $\sigma_{\{u,v\}}$.
	Hence,
	it must pay $\alpha$ for collocating the other pair, in this case $\{u,w\}$,
	or (resp., and) it pays (resp., up to) $\lambda$ for serving (resp., some of) requests in $\sigma_{\{u,w\}}$. 
	Therefore in any case,
	$\OPTM(R_t) \geq  \min{\{ \lambda,\alpha \}}$.
